---
title: "dublinfootfall"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{dublinfootfall}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  fig.width = 7,
  fig.height = 5,
  fig.align = "center",
  message = FALSE,
  collapse = TRUE,
  comment = "#>"
)
```

## Introduction

__dublinfootfall__ is an R package that loads the footfall counts of pedestrians in Dublin in 2021, uses the data to create basic linear regression models, and visualises the data in a variety of plots. Three functions are provided:

1. `load_footfall()`: for loading and tidying the data,

2. `fit_footfall()`: for fitting linear regression models to the data,

3. `visuals()`: for producing visualisations.

These functions should be applied sequentially.

### Installation

__dublinfootfall__ will run in Windows, Mac OS X, or Linux. To install it you first need to install [R](https://cran.r-project.org/). Installing [RStudio](https://www.rstudio.com/) as a nice desktop environment for using R is also recommended.

Once in R you can type at the R command prompt:

```{r, eval=FALSE}
install.packages('devtools')
devtools::install_github('emerphelan/dublinfootfall')
```

to install the latest development version of the package from the __dublinfootfall__ [GitHub page](https://github.com/emerphelan/dublinfootfall). 

In either case, if you then type:

```{r}
library(dublinfootfall)
```

it will load in all the aforementioned __dublinfootfall__ functions.  

## `load_footfall`

This function loads in a data set from Dublin City Council containing footfall counts taken every hour for the duration of 2021. It arranges the data such that it contains 4 columns:   
`Date`: the date the count was taken   
`Time`: the time the count was taken    
`Location`: the name of the location   
`Count`: the footfall count    
There is an option to change the date format, with the options being a completely numeric date, eg. 02-05-2021, or a date with the name of the month, eg. 02-May-2021. In the example below, the format is the former, with the numeric month in the date.

```{r}
dub_dat <- load_footfall("%d-%m-%Y")
head(dub_dat)
```

## `fit_footfall`

This function fits a linear regression model to the data returned by `load_footfall()`. The model uses `Count` as the response with `Date` and `Location` as the predictors. By default, it uses the daily form of the data, but this can be changed to weekly or monthly. It returns an object that contains a variety of different attributes related to the model:  
`mod.residuals`: the residual values  
`mod.fitted`: the fitted values  
`mod.coef`: the model coefficients  
`mod.summary`: the model summary  
`mod.fstat`: the F-statistic  
`mod.call`: the call used to generate the model  
`mod.sigma`: the standard deviation of the random error  
In the example below, a model is generated from the monthly data and the first 10 residual values are printed.  

```{r}
mod <- fit_footfall(dub_dat, data_type = "daily")
mod$mod.residuals[1:10]
```















