---
title: "dublinfootfall"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{dublinfootfall}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  fig.width = 7,
  fig.height = 5,
  fig.align = "center",
  message = FALSE,
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(dublinfootfall)
```

## Introduction

__dublinfootfall__ is an R package that loads the footfall counts of pedestrians in Dublin in 2021, uses the data to create basic linear regression models, and visualises the data in a variety of plots. Three functions are provided:

1. `load_footfall()`: for loading and tidying the data,

2. `fit.footfall()`: for fitting linear regression models to the data,

3. `plot()`: for producing visualisations.

These functions should be applied sequentially.

### Installation

__dublinfootfall__ will run in Windows, Mac OS X, or Linux. To install it you first need to install [R](https://cran.r-project.org/). Installing [RStudio](https://www.rstudio.com/) as a nice desktop environment for using R is also recommended.

Once in R you can type at the R command prompt:

```{r, eval=FALSE}
install.packages('devtools')
devtools::install_github('emerphelan/dublinfootfall')
```

to install the latest development version of the package from the __dublinfootfall__ [GitHub page](https://github.com/emerphelan/dublinfootfall). 

In either case, if you then type:

```{r, eval=FALSE}
library(dublinfootfall)
```

it will load in all the aforementioned __dublinfootfall__ functions.

Some features may not yet be fully tested, and occasionally this version might be liable to break when it is in the process of being updated. 


